Landcover Data Limits Our Understanding of Global Change

### 2015-11-03 Rmd/overview.Rmd
Full draft completed, ready for some edits.

---
title: "Analysis of cropland biases"
author: "Lyndon Estes"
date: "15 November 2014"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes
---

# Introduction

## Sidebar
I am just learning how to use Rmarkdown with Rmd files to document analyses, primarily analyses that are more interactive in nature and notrepeated many times. Code that will be re-used multiple times will go into a library.

One things I am noticing is that the knit process for compiling the notebook does not find functions within code chunks unless the libraries are called within the code chunks themselves. It seems like each code chunk is its own environment.  The solution to this is to do something like `raster::raster()` when you want a function, or run this where chunks are set to {r, eval = FALSE}, and just run the code chunk as if you were in a normal R script working interactively.  I have already set up the library for SAcropland to call the packages its needs so they are loaded up in the global environment.

>>>>

---
title: "Analysis of cropland biases"
author: "Lyndon Estes"
date: "15 November 2014"
output: 
  html_document:
    highlight: tango
    theme: spacelab
---

## Overview

The following files contain the code used to conduct each portion of the analysis described in the accompanying paper __Landcover Data Limits Our Understanding of Global Change__.

#### [Preparing reference maps](cropland-preprocess.html)
#### [Downloading and preparing cropland maps](fetch-land-cover.html)
#### [Cropland map bias and accuracy](compare-landcover.html)
#### [Accuracy versus cropland density](magisterial-district-bias.html)
#### [Carbon bias and accuracy](carbon-bias.html)
#### [Evapotranspiration bias and accuracy](et-bias.html)
#### [Yield and production bias and accuracy](yield-bias.html)
#### [Agent-based model sensitivity](agent-allocation-bias.html)

### 2015-11-03 paper/main.tex
Full draft completed, ready for some edits.

\begin{article}
\begin{abstract}
{Guidelines: landcover maps based on coarser resolution individual sensors (e.g. MODIS) should be aggregated to 50-100 km resolution to minimize map bias and accuracy. Map accuracy is lowest at intermediate levels of cover, thus in highly mixed landscapes, such as small-holder farming systems or ecotones, even coarser aggregation may be necessary. Finer resolutions can be used with Landsat-based maps, if available, or for global analysis from new improved fusion products, particularly GeoWiki. The widely used datasets for carbon, crop area, and crop yield should be rebuilt on these updated datasets. Newer approaches to improving landcover maps, using interactive training with expert judgement and next generation classifiers are needed to create more reliable landcover, and in doing so will improve our understanding of global change. }
\end{abstract}

@@

\dropcap{T}he functioning of the Earth System is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface. Human endeavors are both strongly governed by and shape landcover, whether it be felling ancient forests for timber production or burning savannas to stimulate a green flush for livestock, while landcover is a primary driver of climate and biogeochemical processes \cite{lambin_dynamics_2003}. The vastness of our modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover. To fully understand these processes and the nature of global change, it is therefore essential to know the nature and distribution of landcover.

This importance is understood by a growing number of social, economic, and physical scientists, who increasingly use landcover data to advance understanding of global change in areas ranging from food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010} to carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, and demographic shifts \cite{linard_assessing_2010}.

The value of the insights resulting from such studies depends on the veracity of the landcover data upon which they are based, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations. The reason for this is that landcover data can only practically be derived from satellite imagery, but in many regions the cover types of interest are smaller \cite[e.g. smallholder's farms][]{jain_mapping_2013} than the sensor resolution, or spectrally indistinct from neighboring covers, which translates into substantial mapping errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}. The result is that landcover maps are generally inaccurate at finer scales and disagree substantially with one another, particularly in those parts of the world undergoing the most rapid land use changes \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors mean that we are still unable to obtain the granular, mechanistic understanding of global change processes that we need.

These problems with landcover products are known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}. What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes. Answering this question depends on having spatially comprehensive ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}. Our understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data. This allows us to quantify between map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.

@@

%\vspace{-0.75 cm}
\begin{figure}[!h]
\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scalew.pdf}}
\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map. Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars. Bar colors are coded to specific cropland map, symbols indicate which cover type was used to calculate cropland-adjacent carbon density. The bar represents the mean biases calculated across each of the 5 cover types. Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2). MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
\label{afoto}
\end{figure}

@@

\subsection{Agent-based simulation of food security}
Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}. To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}. In our example, we used an ABM of household food security that simulates the food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}. The model is initialized such that each household is allocated its designated share of cropland, and then produces a crop over the course of one to many seasons. Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g. districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers. To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig. S8) in the eastern part of the country with 28-45\% of their areas devoted to cropland. The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity. The number of households and the cropland area per household is derived from survey data of communities where all cropland is owned. The model is thus considered to well calibrated when all households are allocated their appropriate area of cropland, and all cropland is occupied.

@@

Even more promising are newer approaches that fuse human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power. The GeoWiki project and the new high resolution global forest change map \cite{hansen_high-resolution_2013} provide the most prominent examples of these latest methods. Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \cite{estes_diylandcover:_2015}.  Merging this approach with newer computer vision-based mapping algorithms \cite{debats_generalized_????} is a near-term goal of this project, which will combine the superior capacity of humans to delineate discrete cover types in noisy images \cite{estes_diylandcover:_2015} with the algorithm's speed and ability to classify highly dimensioned image features. These ongoing projects represent the next generation of landcover maps, and offer to greatly improve our understanding of global change.

@@

\begin{materials}
\section{Methods} 
Perhaps it is right {\it SI Materials and Methods}.

Describe weighted mean bias reasons.

We disaggregated the cropland percentages in all maps to binary cropland/non-cropland cover types with 100 m resolution, which matches the typical field size (1 ha) for smallholder farmers in household survey data (collected in Zambia) used in developing the agent-based model \cite{chen_dependency_2013}. The surveys provided the mean cropland area per household (2.2 ha) and frequencies distribution of cropland area holdings across households (e.g. how many households have 1 ha, 2 ha, etc.). We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area, and preserved the cropland area distributions by multiplying the total number of households by the frequencies. We then initialized the model, which takes a weighted (by cropland area frequency) random draw of 100 households and places these within the district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household. This process is iterated until all households are assigned cropland, or all available cropland is allocated.

The South African government commissioned a whole-country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}. The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol. The resulting vectorized field maps, provide a unique, high quality reference dataset describing South African crop field distributions and size classes for the period 2009-2011, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution. We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.

We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates, resulting in 4 maps (hereafter simply the ``cropland maps'') to compare to our reference map.

The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}. The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates. Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.

For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components. In the second example, we examined how these map errors impact the land allocation process of an agent-based food security model \cite{chen_dependency_2013}.

The reference dataset covered all of South Africa's field crop estate, from which we filtered out areas of communally owned farmlands and permanent tree crops to have a common basis of comparison across all landcover products, leaving us with map, after conversion to cropland percentage, that covered or 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.

from agricultural pixels ($>$0.05\% cropland)

The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), as the large uncropped area of South Africa (Fig. 1) dilutes the errors within the

The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}. To be consistent with this methodology, we first adjusted our cropland maps according to this procedure, using the reference map to calculate total cropland area for each of South Africa's nine provinces, then updating the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}. Despite this statistical constraint, the updated cropland maps still had substantial errors that were similar in pattern (Fig. S5) to those in the unadjusted maps (Fig. 1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}. To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.

(which is constrained by an absolute distance and the proximity of other agents)

\section{Digital RCD Analysis} 
\end{materials}

\appendix[App 1]

\appendix
This is an example of an appendix without a title.

\begin{acknowledgments}
I thank everyone tearfully. 
\end{acknowledgments}

>>>>

\begin{article}
\begin{abstract}
{Our knowledge of global change processes is built upon landcover data, but we lack the means to comprehensively assess landcover maps, particularly in rapidly developing regions, and thus do not fully comprehend how map quality impacts our understanding of global change. We used a unique, high quality, national-scale cropland map to assess the bias and accuracy in a) current generation landcover products, and b) analyses built on these. Maps have high bias and low accuracy, particularly at finer resolutions (1-10 km), which transfers into, and is often amplified, in downstream analyses of vegetative carbon stock, gridded crop production, and process-based food security models. Substantial map aggregation is needed to avoid potential misunderstandings of global change processes or misguided policy. Maps derived from coarse resolution sensors (such as MODIS and GlobCover) should be aggregated to at least 50-100 km resolution (and sometimes more). Map accuracy is lowest at intermediate levels of cover, thus coarser aggregation may be necessary in highly mixed landscapes. Higher resolution imaging systems (e.g. Landsat) produce maps that require less aggregation (1-25 km) to achieve high accuracy and low bias. For regional to global analyses, new fusion products, such as the GeoWiki cropland map, need just 5 km of aggregation to produce unbiased maps for most applications, but can require $\geq$25 km of aggregation to have sufficient accuracy. Widely used global datasets for carbon, crop area, and crop yield should be rebuilt on GeoWiki and similar maps, while new approaches methods that fuse human judgement, the latest computer vision-based algorithms, and growing high resolution image archives are needed to produce the next generation of landcover maps. }
\end{abstract}

@@

\dropcap{T}he functioning of the Earth System is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface. Human endeavors are strongly governed by and shape landcover, whether it be felling forests for timber or burning savannas to rejuvenate forage, and landcover is a primary factor in many climate and biogeochemical processes \cite{lambin_dynamics_2003}. Our increasing modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover. To fully understand these processes and the nature of global change, it is essential to know the nature and distribution of landcover.

This importance is understood by a growing number of social, economic, and natural scientists, who are using landcover data to advance understanding of food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, demographic shifts \cite{linard_assessing_2010}, and other important facets of global change.

The value of the insights resulting from such studies depends upon the veracity of their underlying landcover data, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations. The reason for this is that landcover data can only practically be derived from satellite imagery, and in many regions the cover types of interest are smaller \cite[e.g. smallholder's farms][]{jain_mapping_2013} than the sensor resolution, or spectrally indistinct from neighboring covers, which translates into substantial mapping errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}. Landcover maps are therefore often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors mean that we are still unable to obtain the granular, mechanistic understanding of global change processes that we need.

These problems with landcover products are known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}. What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes. Answering this question depends on having spatially comprehensive ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}. Our understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data. This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.

@@

%\vspace{-0.75 cm}
\begin{figure}[!h]
\centerline{\includegraphics[width=.5\textwidth]{figures/figure3.pdf}}
\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map. Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars. Bar colors are coded to specific cropland map, symbols indicate which cover type was used to calculate cropland-adjacent carbon density. The bar represents the mean biases calculated across each of the 5 cover types. Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2). MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
\label{afoto}
\end{figure}

@@

\subsection{Agent-based simulation of food security}
Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}. To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}. In our example, we used an ABM of household food security that simulates the food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}. The model is initialized such that each household is allocated its designated share of cropland, and then produces a crop over the course of one to many seasons. Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g. districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers. To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig. S8) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland. The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity. The number of households and the cropland area per household was derived from surveys of farming communities where all cropland is used. The model was thus considered to well calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated.

@@

Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of landcover maps. The GeoWiki project and the new high resolution global forest change map \cite{hansen_high-resolution_2013} are the most advanced examples of these newer methods. Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \cite{estes_diylandcover:_2015}. Merging this crowdsourced approach to vector map creation with newer computer vision-based mapping algorithms \cite{debats_generalized_????} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \cite{estes_diylandcover:_2015}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets. Approaches such as these will be needed to provide the next generation of landcover maps, so that we can gain a clearer, unbiased understanding of global change.

@@

\begin{materials}
\section{Methods} 
To establish the accuracy of the cropland boundary reference map, we randomly selected 609 cell from a 1 km$^2$ grid covering South Africa, weighted by the probability of cropland occurrence \cite{estes_diylandcover:_2015}. We sub-divided each cells into 200 m sub-cells, and visually interpreted whether crop fields presence/absence within these against Google Earth imagery, and compared these results to the digitized field polygons to determine the reference map's sensitivity (0.89) and specificity (0.99), overall map accuracy (97\%), and its true skill statistic \cite[0.88;][]{allouche_assessing_2006}. We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map. In the resulting grid, we masked out areas classified as communal farmland, because only their outer perimeters were digitized \cite{fourie_better_2009}, which posed a risk of overestimating cropland extent. We also excluded the permanent tree crop class, to avoid confusion with landcover maps that include these in their cropland classes. For the same reason, we used other high resolution landcover maps for selected provinces of South Africa to mask out commercially afforested areas and sugarcane plantations, which were also not included in the reference map. The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.
We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution (GeoWiki is already a cropland percentage map). Both MODIS and GlobCover had mixed pixel classes, thus we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages. We used the mean map for the main analysis, but estimated error variability using all three versions (SI).\\  
\indent To uncover the relationship between map error and cover density, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$, see SI) to provide a landscape-scaled unit for calculating characteristic cover density. We filtered out pixels with $<$0.05\% (0.5 ha) cropland, to prevent the much larger areas of non-agricultural districts from drowning out any relationship. We extracted the absolute values of test map errors and corresponding reference cropland percentages, calculated their district-wide means, and then modeled the relationship between mean absolute error (response) and cropland density (predictor) as a polynomial function within a generalized additive model. We accounted for potential spatial autocorrelation by fitting a two-dimensional smoothing spline to each district's centroid coordinates, and weighted the model by the number of agricultural pixels per district \cite{wood_mgcv:_2001}.\\   
\indent To develop the carbon maps, we followed the methods used by Ruesch and Gibbs \cite{ruesch_new_2008} to create their vegetative carbon map. We took the average of the biome specific carbon density values they provided for African forests, secondary forests, shrublands, grasslands, sparse habitats, and croplands. We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps. We aggregated each carbon map up to the five coarser resolutions for scaling comparisons.  \\
\indent We used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model run with the Africa Flood and Drought Monitor's meteorological data \cite{sheffield_drought_2013} to produce monthly gridded evapotranspiration estimates for South Africa for the years 1979-2010. We adjusted VIC's native landcover scheme so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25${\circ}$ resolution). We ran one instance of VIC for each of the five landcover permutations, and compared the mean annual ET produced by the reference map variant with those from the test maps. \\
\indent To create the gridded maize yield and production maps, we followed \cite{ramankutty_farming_2008} to first calibrate the test cropland maps against cropland areas reported by administrative-level agricultural censuses. In place of census reports, we used the reference map to calculate total cropland areas for South Africa's nine provinces, then adjusted the pixel-wise cropland percentages in the four test maps so that their province-wise sums matched these totals (SI). We then followed Monfreda et al's procedure \cite{monfreda_farming_2008} to disaggregate maize \cite[South Africa's largest crop][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps. Per \cite{monfreda_farming_2008}, the gridded planted area and yield estimates were obtained from magisterial district-level agricultural censuses, in this case for the most recent census of 2007 \cite{statistics_south_africa_commercial_2007}. We then aggregated gridded harvested areas and yields (weighted by the corresponding harvested area) to each coarser resolution, and multiplied yield by harvested area to calculate production.\\
\indent The agent-based model of food security used here calculates average household-level food production over multiple seasons based on the biophysical environment (soil properties, weather), management inputs (planting decisions, labor availability), and farm assets (total field area) \cite{chen_dependency_2013}. The frequency distributions of household cropland holdings were derived from household survey data obtained from Zambia (no equivalent data were available for South Africa). We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies. To create cropland surfaces for the, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha. In initializing, the model takes a weighted (by cropland area frequency) random draw of 100 households and places these within a district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household. This process is iterated until all households are assigned cropland, or all available cropland is allocated. The model then runs separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions. We ran the model for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).  \\
\indent In calculating map error statistics, and with the exception of the agent-based model results, we use the reference cropland percentages to calculate the weighted mean map error (bias) and mean absolute map error (accuracy). For the comparisons between aggregated map values, we applied a further weight--the number of pixels contributing to each aggregated pixel, to prevent pixels close to national boundaries or where non-target cover types were masked out from having outsize influence on the statistics.

%The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}. The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates. Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.

%The reference dataset covered all of South Africa's field crop estate, from which we filtered out areas of communally owned farmlands and permanent tree crops to have a common basis of comparison across all landcover products, leaving us with map, after conversion to cropland percentage, that covered or 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.

%The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), as the large uncropped area of South Africa (Fig. 1) dilutes the errors within the 
%(which is constrained by an absolute distance and the proximity of other agents)

%\section{Digital RCD Analysis} 
\end{materials}

%\appendix[App 1]

%\appendix
%This is an example of an appendix without a title.

\begin{acknowledgments}
This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program
(NNX15AC64G), and the National Science Foundation (SES-1360463 and BCS-1026776).
\end{acknowledgments}
